Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-08-02 analysis/paper/hotspot-one-chromosome.Rmd
replaced hot_chr with {{hot_chr}}

```{r, def-{{hot_chr}}}
hot_lower <- {{hot_mid}} - 2
hot_upper <- {{hot_mid}} + 2
fp <- paste0("Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
```

```{r, readRDS-{{hot_chr}}}
pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
```

```{r, define_local_annot-{{hot_chr}}}
local_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r, define_nonlocal_annot-{{hot_chr}}}
nonlocal_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
```

@@

```{r, scatter-{{hot_chr}}}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == {{keller_mediator}}, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
plotly::ggplotly(p)
```

@@

fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
plotly::ggplotly(p)
```

@@

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
}
```

@@

scols_pre <- scols_extra %>%
  dplyr::select(n_med_per_local$local_symbol)
# get ordering for columns of scols_pre
col_order <- scols_pre %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
  order.dendrogram()
scols <- scols_pre[ , col_order] %>%
  mutate(local = scols1)
fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
```

@@

(p <- pm[ord, ord] %>%
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
)
fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
htmlwidgets::saveWidget(widget = p, 
                        file = fnhtml
                        )
```

>>>>

```{r, def-{{hot_chr}}}
hot_lower <- {{hot_mid}} - 2
hot_upper <- {{hot_mid}} + 2
fp <- paste0("Chr", {{hot_chr}}, "-")
knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
```

```{r, readRDS-{{hot_chr}}}
pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio.rds")))
pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_local.rds")))
pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_nonlocal.rds")))
med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_mediations.rds")))
```

```{r, define_local_annot-{{hot_chr}}}
local_annot <- lod_peaks %>%
  filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == {{hot_chr}}) %>%
  filter(local_tx)
```

```{r, define_nonlocal_annot-{{hot_chr}}}
nonlocal_annot <- lod_peaks %>%
  filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == {{hot_chr}}) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
```

@@

```{r, scatter-{{hot_chr}}}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == {{keller_mediator}}, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter.svg")
fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
plotly::ggplotly(p)
```

@@

fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter-panel.svg")
fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter-panel.pdf")
ggsave(p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
plotly::ggplotly(p)
```

@@

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter-panel", i, ".pdf")
  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
}
```

@@

scols_pre <- scols_extra %>%
  dplyr::select(n_med_per_local$local_symbol)
# get ordering for columns of scols_pre
col_order <- scols_pre %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
  order.dendrogram()
scols <- scols_pre[ , col_order] %>%
  mutate(local = scols1)
fnhtml <- here("analysis", "figures", paste0("Chr", {{hot_chr}}, "hm.html"))
```

@@

(p <- pm[ord, ord] %>%
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
)
fnpdf <- here("analysis", "figures", paste0("Chr", {{hot_chr}}, "hm.pdf"))
htmlwidgets::saveWidget(widget = p, 
                        file = fnhtml
                        )
```

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
replaced hot_chr with {{hot_chr}}

---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "7/27/2019"
bibliography: research.bib
linkReferences: true
link-citations: true
output: 
  - bookdown::html_document2:
    code_folding: hide
  - bookdown::pdf_document2:
    code_folding: hide  
---

@@

```{r setup}
library(qtl2hotspots)
pleio_threshold <- 4
ult <- 15 # univariate lod threshold
```

@@

```{r setup_continued-2}
hot_chr <- 2
hot_mid <- 165.5
keller_mediator <- "Hnf4a"
out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
```

@@

@@

```{r setup_continued-5}
hot_chr <- 5
hot_mid <- 146
keller_mediator <- "Pdx1"
out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
```

>>>>

---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "7/27/2019"
bibliography: research.bib
linkReferences: true
link-citations: true 
output: 
  bookdown::html_document2:
    code_folding: hide
---

@@

```{r setup}
library(knitr)
library(qtl2hotspots)
pleio_threshold <- 4
ult <- 15 # univariate lod threshold
```

@@

```{r, load_purrr}
library(purrr)
```

```{r setup_continued}
hot_chr <- c(2, 5, 7, 11, 13)
hot_mid <- c(165.5, 146, 46, 71, 112.5)
keller_mediator <- c("Hnf4a", "Pdx1", "Fam83e", "Sat2", "Il6st")
out <- pmap(tibble(hot_chr, hot_mid, keller_mediator), function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))))
```

@@

`r knitr::knit(out[[1]])`

@@

```{r setup_continued-5}
hot_chr <- 5
hot_mid <- 146
keller_mediator <- "Pdx1"
out <- knitr::knit_expand(file = "hotspot-one-chromosome.Rmd")
```
